---
subtitle: "{{< var course_title >}}"
title: "Descriptive statistics"
author: Natapol Pornputtapong, PhD.
institute: '[Chulalongkorn University](#)'
toc: true
toc-depth: 1
toc-title: "What we will cover today:"
center-title-slide: false
from: markdown+emoji
date: 20 October 2023
date-meta: 20 October 2023
date-format: "DD MMM YYYY"
editor: 
  markdown: 
    wrap: 72
format:
  revealjs:
    fig-responsive: true
    theme: simple
    scrollable: true
    echo: true
    slide-number: c/t
    mouse-wheel: false
    preview-links: auto
    logo: /figures/icons/course_favicon.png
    footer: "{{< var course_title >}}"
---

# PPDAC: A for Analysis

# Descriptive statistics

## library installation

- tidyverse
  - tibble
  - tribble
  - readr
  - tidyr
  - dplyr
  - ...
- tidymodels
  - infer
  
```{r}
#| eval: false
install.packages("tidyverse")
install.packages("tidymodels")
```

## calling library

```{r}
#| code-fold: true
#| code-summary: "Show the code"

library(tidyverse)
library(tidymodels)
```

## The NHANES data

| Variable           | Definition                                                                                                                                                                               |
|:--------------------------------|:--------------------------------------|
| id                 | A unique sample identifier                                                                                                                                                               |
| Gender             | Gender (sex) of study participant coded as male or female                                                                                                                                |
| Age                | Age in years at screening of study participant. Note: Subjects 80 years or older were recorded as 80.                                                                                    |
| Race               | Reported race of study participant, including non-Hispanic Asian category: Mexican, Hispanic, White, Black, Asian, or Other. Not availale for 2009-10.                                   |
| Education          | Educational level of study participant Reported for participants aged 20 years or older. One of 8thGrade, 9-11thGrade, HighSchool, SomeCollege, or CollegeGrad.                          |
| MaritalStatus      | Marital status of study participant. Reported for participants aged 20 years or older. One of Married, Widowed, Divorced, Separated, NeverMarried, or LivePartner (living with partner). |
| RelationshipStatus | Simplification of MaritalStatus, coded as Committed if MaritalStatus is Married or LivePartner, and Single otherwise.                                                                    |
| Insured            | Indicates whether the individual is covered by health insurance.                                                                                                                         |
| Income             | Numerical version of HHIncome derived from the middle income in each category                                                                                                            |
| Poverty            | A ratio of family income to poverty guidelines. Smaller numbers indicate more poverty                                                                                                    |
| HomeRooms          | How many rooms are in home of study participant (counting kitchen but not bathroom). 13 rooms = 13 or more rooms.                                                                        |
| HomeOwn            | One of Home, Rent, or Other indicating whether the home of study participant or someone in their family is owned, rented or occupied by some other arrangement.                          |
| Work               | Indicates whether the individual is current working or not.                                                                                                                              |
| Weight             | Weight in kg                                                                                                                                                                             |
| Height             | Standing height in cm. Reported for participants aged 2 years or older.                                                                                                                  |
| BMI                | Body mass index (weight/height2 in kg/m2). Reported for participants aged 2 years or older.                                                                                              |
| Pulse              | 60 second pulse rate                                                                                                                                                                     |
| BPSys              | Combined systolic blood pressure reading, following the procedure outlined for BPXSAR.                                                                                                   |
| BPDia              | Combined diastolic blood pressure reading, following the procedure outlined for BPXDAR.                                                                                                  |
| Testosterone       | Testerone total (ng/dL). Reported for participants aged 6 years or older. Not available for 2009-2010.                                                                                   |
| HDLChol            | Direct HDL cholesterol in mmol/L. Reported for participants aged 6 years or older.                                                                                                       |
| TotChol            | Total HDL cholesterol in mmol/L. Reported for participants aged 6 years or older.                                                                                                        |
| Diabetes           | Study participant told by a doctor or health professional that they have diabetes. Reported for participants aged 1 year or older as Yes or No.                                          |
| DiabetesAge        | Age of study participant when first told they had diabetes. Reported for participants aged 1 year or older.                                                                              |
| nPregnancies       | How many times participant has been pregnant. Reported for female participants aged 20 years or older.                                                                                   |
| nBabies            | How many of participants deliveries resulted in live births. Reported for female participants aged 20 years or older.                                                                    |
| SleepHrsNight      | Self-reported number of hours study participant usually gets at night on weekdays or workdays. Reported for participants aged 16 years and older.                                        |
| PhysActive         | Participant does moderate or vigorous-intensity sports, fitness or recreational activities (Yes or No). Reported for participants 12 years or older.                                     |
| PhysActiveDays     | Number of days in a typical week that participant does moderate or vigorous-intensity activity. Reported for participants 12 years or older.                                             |
| AlcoholDay         | Average number of drinks consumed on days that participant drank alcoholic beverages. Reported for participants aged 18 years or older.                                                  |
| AlcoholYear        | Estimated number of days over the past year that participant drank alcoholic beverages. Reported for participants aged 18 years or older.                                                |
| SmokingStatus      | Smoking status: Current Former or Never.                                                                                                                                                 |

## import data to tibbles

```{r}

nh <- read_csv(file="nhanes.csv")
nh
```

## view tibbles

```{r}
#| eval: false

View(nh)
```

## Display all Race values

```{r}

nh$Race
```

## Get the unique values of Race

```{r}
nh$Race %>% unique
nh$Race %>% unique %>% length
```

## Age mean, median, range

```{r}
nh$Age %>% mean
nh$Age %>% median
nh$Age %>% range
```

# exploratory data analysis (EDA)

## Histogram

```{r}

nh %>% 
  ggplot(aes(BMI)) + 
  geom_histogram(bins=30)
```

## Histogram with calculation

```{r}

# In pounds, more bins

nh %>% 
  ggplot(aes(Weight*2.2)) + 
  geom_histogram(bins=80)
```

## Scatter plots

```{r}

nh %>% 
  ggplot(aes(Height, Weight, col=Gender)) + 
  geom_point()
```

## Complex scatter plots

```{r}

nh %>% 
  filter(Age>=18) %>% 
  ggplot(aes(Height, Weight, col=Gender)) + 
    geom_point() + 
    geom_smooth(method="lm")
```

## References

-   https://sparkbyexamples.com/r-programming/select-rows-in-r/
-   https://4va.github.io/biodatasci/
-   https://bioconnector.github.io/
